---
title: "Homework 7"
author: "Sarah Morris"
date: "3/17/2021"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---  

#### Part 1: Fake Data

```{r, echo=TRUE}

library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation

```


##### Generate fake data
```{r, echo=TRUE}
# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)
```

##### plot histogram
```{r, echo=TRUE}
# ..density.. --> shorthand for ggplot to plot the density
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)

```

##### add empirical density curve
```{r, echo=TRUE}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

##### get maximum likelihood parameters for normal distribution
```{r, echo=TRUE}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```

##### plot normal probability density 
###### call the dnorm function inside ggplot’s stat_function to generate the probability density for the normal distribution
###### we first get the maximum likelihood parameters for a normal distribution fitted to thse data by calling fitdistr. Then we pass those parameters (meanML and sdML to stat_function
```{r, echo=TRUE}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML)) #red line = probability density
 p1 + stat
 
# the best-fitting normal distribution (red curve) for these data actually has a biased mean. That is because the data set has no negative values, so the normal distribution (which is symmetric) is not working well
```

##### plot exponential probability density
```{r, echo=TRUE}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
```

##### plot uniform probability density
```{r, echo=TRUE}
# For the uniform, we don’t need to use fitdistr because the maximum likelihood estimators of the two parameters are just the minimum and the maximum of the data
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
```

##### plot gamma probability density 
```{r, echo=TRUE}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="purple", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

##### plot beta probability density
```{r, echo=TRUE}

#This one has to be shown in its own plot because the raw data must be rescaled so they are between 0 and 1, and then they can be compared to the beta

pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),
                     "beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, 
                             colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```
 

#### Part 2: Actual Data


```{r, echo=TRUE}
bacteria <- read.csv("Tuesday_ZOI.csv")
as.data.frame(bacteria)
str(bacteria)
summary(bacteria)

#z <- rnorm(n=3000,mean=0.2)
#z <- data.frame(1:3000,z)
#names(z) <- list("ID","myVar")
#z <- z[z$myVar>0,]
#str(z)
#summary(z$myVar)

p1 <- ggplot(data=bacteria, aes(x=bacteria$width, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)

p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

normPars <- fitdistr(bacteria$width,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute

meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(bacteria$width),len=length(bacteria$width))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(bacteria$width), args = list(mean = meanML, sd = sdML))
 p1 + stat
 
expoPars <- fitdistr(bacteria$width,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(bacteria$width), args = list(rate=rateML))
 p1 + stat + stat2
 
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, 
                       colour="darkgreen", n = length(bacteria$width), args = list(min=min(bacteria$width), max=max(bacteria$width)))
 p1 + stat + stat2 + stat3
 
gammaPars <- fitdistr(bacteria$width,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]
print(gammaPars)

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="purple",
                       n = length(bacteria$width), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
 
pSpecial <- ggplot(data=bacteria, aes(x=width/(max(width + 0.1)), 
                                      y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=bacteria$width/max(bacteria$width + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta,
                             colour="orchid", n = length(bacteria$width), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```

###### Okay, so none of the probability distributions fit great
###### which is probably because there are 2 different strains of bacteria in this dataset
###### next time need to split apart the different strains into 2 datsets
###### but of all the distributions, I guess the gamma fits the best

###### maximum likelihood parameters for gamma distribution
shape = 5.743
rate = 0.277

```{r, echo=TRUE}

#from before
gammaPars <- fitdistr(bacteria$width,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]
print(gammaPars)

z2 <- rgamma(n=35,shape = 5.743, scale = 0.277)
z2 <- data.frame(1:35,z2)
names(z2) <- list("ID","myVar")
z2 <- z2[z2$myVar>0,]
str(z2)
summary(z2$myVar)

xval2 <- seq(0,max(z2$myVar),len=length(z2$myVar))

p3 <- ggplot(data=z2, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p3)

gammaPars <- fitdistr(z2$myVar,"gamma")
shapeML2 <- gammaPars$estimate["shape"]
rateML2 <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval2, y = ..y..), fun = dgamma, colour="brown", 
              n = length(z2$myVar), args = list(shape=shapeML2, rate=rateML2))
 p3 + stat4

p2 <- ggplot(data=bacteria, aes(x=width, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p2)

stat5 <- stat_function(aes(x = xval2, y = ..y..), fun = dgamma, colour="purple", 
          n = length(bacteria$width), args = list(shape=shapeML, rate=rateML))
 p2 + stat5

````

###### Yes, the model is doing a good job of simulating realistic data; however the x-axis is a power of 10 smaller than the original data. These data are hard to fit a curve to, so given that, I think the simulation is as good as it can be. Perhaps if my original data fit a gamma distribution more closely, the simualtion would be even more realistic. I'm also wondering if there is a numerical measurement to see how well the data fit the distribution. 
  
  
