---
title: "Homework 11"
author: "Sarah Morris"
date: "4/21/2021"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

#### Batch Processing

```{r}
# using real data to simulate multiple files of data for batch processing
sporedata <- read.csv("spore_size.csv")

sporedata[] <- lapply(sporedata, as.numeric)

clean_spores <- sporedata[complete.cases(sporedata),] # subset non-NA data/clean cases

length_avg <- mean(clean_spores$Length)
width_avg <- mean(clean_spores$Width)
vol_avg <- mean(clean_spores$Volume)

lengthSD <- sd(clean_spores$Length)
widthSD <- sd(clean_spores$Width)
volSD <- sd(clean_spores$Volume)

# V <- V[sign(V) == 1] # to eliminate negative values

####################################################################
# ------------------------------------------------------
# FUNCTION file_builder
# description: creates a set of random files for regression
# inputs: file_n = number of files to create 
#         file_folder = name of folder for random files
#         file_size = c(min,max) number of rows in file
# outputs: creates a set of random files
########################################################
file_builder <- function(file_n=10,
                         file_folder="SporeFiles/", 
                         file_size=c(15,100)) 
                          { 
  for (i in seq_len(file_n)) { 
    file_length <- sample(file_size[1]:file_size[2], size=1) 
    spore_length <- rnorm(n=file_length, mean=length_avg, sd=lengthSD) 
    spore_width <- rnorm(n=file_length, mean=width_avg, sd=widthSD) 
    . <- rnorm(n=file_length, mean=vol_avg, sd=volSD) 
    spore_vol <- replace(., which(. < 0), NA) # replace negative values with NAs
    df <- data.frame(spore_length,spore_width,spore_vol) 
    
# create label for file name with padded zeros
    file_label <- paste(file_folder, 
                        "Spore",
                        formatC(i, # takes formatting from C language
                                width=3, 
                                format="d",
                                flag="0"),
                        ".csv",sep="")
# set up data file and incorporate time stampe and minimal metadata
    write.table(cat("# Simulated spore data file for batch processing", "\n",
                    "# timestamp: ",as.character(Sys.time()), 
                    "\n", "# SKM", "\n",
                    "# ----------------------------------", "\n",
                    "\n",
                    file=file_label, 
                    row.names="",
                    col.names="",
                    sep=""))
    
# now add the data frame
    write.table(x=df, 
                file=file_label, 
                sep=",", 
                row.names=FALSE, 
                append=TRUE) 
  } # end of for loop
    
} # end of file_builder
# ------------------------------------------------------

# ------------------------------------------------------
# FUNCTION reg_stats
# description: fits linear model, extract stats
# inputs: 3 column dataframe   
# outputs: slope,p-value, and r^2 
########################################################
reg_stats <- function(d=NULL) {
             if(is.null(d)) {
               len <- runif(10)
               wid <- runif(10) # going to ignore this for simplicity
               vol <- runif(10)
               d <- data.frame(len,wid,vol)
             }
  . <- lm(data=d,d[,3]~d[,1])
  . <- summary(.)
  VxL_stats_list <- list(Slope=.$coefficients[2,1],
                     pVal=.$coefficients[2,4],
                     r2=.$r.squared)
  . <- lm(data=d,d[,3]~d[,2])
  . <- summary(.)
  VxW_stats_list <- list(Slope=.$coefficients[2,1],
                     pVal=.$coefficients[2,4],
                     r2=.$r.squared)
  
  
  return(c(VxL_stats_list,VxW_stats_list))
  
} # end of reg_stats
# ------------------------------------------------------

library(TeachingDemos)
char2seed("Flatpicking solo")

########################################################
# Global Variables
file_folder <- "SporeFiles/"
n_files <- 50
file_out <- "Spore_Stats.csv"
########################################################

# create data sets
dir.create(file_folder)
file_builder(file_n=n_files) #col names warning is ok to ignore
file_names <- list.files(path=file_folder)

# create a data frame to hold summary file statistics
ID <- seq_along(file_names)
file_name <- file_names
VxL_slope <- rep(NA, length(file_names))
VxL_pVal <- rep(NA, length(file_names))  
VxL_r2 <- rep(NA, length(file_names))  
VxW_slope <- rep(NA, length(file_names))
VxW_pVal <- rep(NA, length(file_names))  
VxW_r2 <- rep(NA, length(file_names))  

stats_out <- data.frame(ID,file_name,VxL_slope,VxL_pVal,VxL_r2,
                        VxW_slope,VxW_pVal,VxW_r2)

# batch process by looping by looping through individual files

for (i in seq_along(file_names)) { # to get current # of files in the data set
     data <- read.table(file=paste(file_folder,file_names[i],sep=""),
                        sep=",",
                        header=TRUE)
     d_clean <- data[complete.cases(data),] # subset non-NA data/clean cases
     
     . <- reg_stats(d_clean) # pull out regression stats from clean file
     
     stats_out[i,3:8] <- unlist(.) 
}

# set up output file and incorporate time stamp and minimal metadata

# write.table(cat("# Summary stats for ", 
#                 "batch processing of regression models",
#                 "\n",
#                 "# timestamp: ", as.character(Sys.time()),
#                 "\n",
#                 file=file_out,
#                 row.names="",
#                 col.names="",
#                 sep=""))

# now add the data frame
write.table(x=stats_out,
            file=file_out,
            row.names=FALSE,
            col.names=TRUE,
            sep=",",
            append=TRUE) # don't worry about warning message re appending col names

```
Yay! It worked. I created 50 files in Spore files with spore width, length and volume measurements, and I created a summary .csv with the calculates slopes, p-values and r squared values from the linear regression models of 1) length and volume and 2) width and volume.  


Just for kicks, I'm going to calculate the average summary stats
```{r}
data <- read.csv("Spore_Stats.csv")
print(sapply(data[3:8],mean))
```
Looks like the relationship isn't all that strong between length and volume or width and volume...  
