---
title: "Homework 12"
author: "Sarah Morris"
date: "4/28/2021"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

#### ggplot  

I don't have good datasets for this, so I will use one of the built in datasets.
I've chosen the **msleep** dataset, which contains sleep data for mammals.

variables in the dataset:  

  * name = common name  
  * genus  
  * vore = carnivore/omnivore/herbivore  
  * order  
  * conservation = conservation status of the animal  
  * sleep_total = total sleep, in hours  
  * sleep_rem = rem sleep, in hours  
  * sleep_cycle = length of slep cycle, in hours  
  * awake = amount of time spent awake, in hours  
  * brainwt = brain weight in kilograms  
  * bodywt = body weight in kilograms  


```{r setup, warning=FALSE} 

# Preliminaries
library(ggplot2)
library(ggthemes)
library(patchwork)
library(colorblindr) # requires colorspace
library(cowplot)
library(colorspace) 
library(ggsci)


d <- msleep
# exploring msleep
str(msleep)
# vore is 3 distinct categories --> good for bar and boxplots

d <- subset(d, select = -c(sleep_cycle,sleep_rem)) # removing these cause they have loads of NAs
#d <- d[complete.cases(d),] # subset non-NA

# first question: does brain weight predict body weight?
# scatterplot of brainwt by sleep_total split into 3 vore categories by colors
p1 <- ggplot(data=d,mapping=aes(x=brainwt,y=bodywt, color=vore, fill=vore)) +
geom_point() + geom_smooth(level=0.80)
print(p1)
# I suspect elephants are an outlier here, so I'm going to remove them and redo the analysis

e <- d[-c(21,36),] # elephant-less dataset

p2 <- ggplot(data = e, mapping = aes(x=e$brainwt,y=e$bodywt, color=e$vore, fill=e$vore)) +
geom_point(size=3)+
geom_smooth(level=0.80)+
  xlab("brain weight")+
  ylab("body weight")+
  ggtitle("Mammal body weight x brain weight", subtitle = "divided by diet type")+
  theme_dark() + theme(
    panel.background = element_rect(fill = "#2D2D2D"),
    axis.title = element_text(size = 20),
    plot.title = element_text(size = 25),
    plot.subtitle = element_text(size = 15),
    axis.text = element_text(size = 15),
    legend.key = element_rect(fill = "#2D2D2D")) 
  
print(p2) # this is a slightly better looking plot

# ggsave(filename="MyScatterPlot.jpg",plot=p2,
#        device="jpg",width=20,height=20,units="cm",dpi=300)

```


Next question: does dietary type predict total sleep time?

```{r, warning=FALSE} 

sum(is.na(d$vore)) # 7 NAs
which(is.na(d$vore)) # identifies rows with NA for 'vore'
f <- d[-which(is.na(d$vore)),] # vore NA-less dataset

p3 <- ggplot(f,aes(x=vore,y=sleep_total,fill=vore)) +
  geom_boxplot(outlier.shape=NA) + 
  geom_point(position=position_jitter(width=0.1,height=0.7),color="grey40",size=2)

print(p3)

library(scales)
mypal <- pal_rickandmorty("schwifty", alpha = 0.9)(12)
mypal
show_col(mypal)

p3 + scale_fill_manual(values=mypal[c(1,3,4,6)]) +
annotate("segment", x = 2, xend = 2.9, y = 25, yend = 18.5, colour = "red", size=3, alpha=0.6, arrow=arrow())
```
Insectivores sleep the most!  


Finally, I want to see how sleep total varies by dietary type across conservation statuses

```{r, warning=FALSE} 

# new facet label for conservation status
con_lab <- as_labeller(c(`cd`="cd",`domesticated`="domesticated",`en`="endangered",`lc`="least concern", `nt`="not threatened", `vu`="vulnerable"),multi_line = TRUE)

# do a facet with 3 groups - brainwt, sleep_total & vore and conservation
m1 <- ggplot(data=f, mapping=(aes(x=vore,y=sleep_total,fill=vore))) +
  geom_boxplot() 
m1 + facet_grid(.~conservation,
  labeller = labeller(conservation=con_lab))

```

Unsuprisingly, there isn't loads of data for endangered or vulnerable species.   

