---
title: "Homework 8"
author: "Sarah Morris"
date: "3/31/2021"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

#### Homework 8: Creating Fake Data Sets to Explore Hypotheses

##### I will be using data from this publication:  
**Li, M., Zheng, Y., Fan, R., Zhong, Q., & Cheng, D. (2017).** Scaling relationships of twig biomass allocation in Pinus hwangshanensis along an altitudinal gradient. PLOS ONE, 12(5). 
  
The means, variances and sample sizes for low and high elevational groups come from this meta-analysis that included the above paper:   
**Midolo G, De Frenne P, Hölzel N, Wellstein C. 2019.** Global patterns of intraspecific leaf trait responses to elevation. Global Change Biology 25: 2485–2498.

##### The **hypothesis** I am testing =  
Within species with broad elevational ranges, individuals found at higher elevations will have denser leaves (higher mass per unit leaf area (**LMA**)). Alpine plant species tend to be smaller and more compact, so I believe that phenotypic plasticity within species will also trend the same way.

```{r}
data <- read.csv("Pinus.csv")

str(data) # checking structure and column label names

# renaming columns
data$LMA <- data$X.LMA..mg.mm2. # LMA = leaf mass per area (mg/mm^2)
data$Alt <- data$X.Alt # Alt = elevation (m)

# the LMA values are annoyingly small, so I'm converting the measurement to g/mm^2
data$LMA <- data$LMA*1000
data$LMA # these numbers are easier to look at


low <- subset(data, Alt == "1200") # low elevation = 1200m
mid <- subset(data, Alt == "1600") # mid elevation = 1600m
high <- subset(data, Alt == "2000") # high elevation = 2000m 

# expected mean LMA for low elevation = 100 g/mm^2
# expected mean LMA for mid elevation = 150 g/mm^2
# expected mean LMA for high elevation = 200 g/mm^2
# standard deviation = 10
# sample size for each group = 45

# making the random data set
nGroup <- 3 # number of treatment groups
nName <- c("low","mid","high") # names of groups
nSize <- c(45,45,45) # number of observations in each group
nMean <- c(100,150,200) # expected mean of each group
nSD <- c(10,10,10) # standard deviation of each group
ID <- 1:(sum(nSize)) # assigning and id number for each row

# putting the data into a dataframe
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]))

TGroup <- factor(x=rep(nName,nSize), levels=nName, ordered= is.ordered(nName)) # make a column with the nNames repeated nSize times as factor data so that order of elevational groups stays the way I want

ANOVAdata <- data.frame(ID,TGroup,resVar) # putting it all together

str(ANOVAdata) # checking it out
hist(ANOVAdata$resVar[TGroup == "low"])
hist(ANOVAdata$resVar[TGroup == "mid"]) #less normally distributed than I'd expect
hist(ANOVAdata$resVar[TGroup == "high"])

```

##### Question 4: analyze the data   
ANOVA + boxplot
```{r}
ANOVAmodel <- aov(resVar~TGroup,data=ANOVAdata) # ANOVA analysis
boop <- summary(ANOVAmodel) # put the summary into an object
boop # take a peek
aggregate(resVar~TGroup,data=ANOVAdata,FUN=mean) # print the means for low, mid, high
unlist(boop) # to make it easier to extract the ones we want
ANOVAsum <- list(Fval=unlist(boop)[7],probF=unlist(boop)[9]) # F statistic & p-value
ANOVAsum
```
P-value of the ANOVA is *very* significant  
Now, the boxplot:  
```{r}
library(ggplot2) # load ggplot

ANOVAplot <- ggplot(data=ANOVAdata,aes(x=TGroup,y=resVar,fill=TGroup)) +
  geom_boxplot() +
  xlab("elevational group") +
  ylab("leaf density")

print(ANOVAplot)

ggsave(filename = "ANOVA_HW8.pdf", plot = ANOVAplot, device = "pdf") #in case I want to look at this later
```

##### Question 5: running the analysis multiple times  
copying streamlined code down below  
```{r}
nGroup <- 3 
nName <- c("low","mid","high")
nSize <- c(45,45,45) 
nMean <- c(100,150,200)
nSD <- c(10,10,10) 
ID <- 1:(sum(nSize)) 
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]))
TGroup <- factor(x=rep(nName,nSize), levels=nName, ordered= is.ordered(nName)) 
ANOVAdata <- data.frame(ID,TGroup,resVar) 
ANOVAmodel <- aov(resVar~TGroup,data=ANOVAdata) 
boop <- summary(ANOVAmodel) 
aggregate(resVar~TGroup,data=ANOVAdata,FUN=mean)
unlist(boop)
ANOVAsum <- list(Fval=unlist(boop)[7],probF=unlist(boop)[9])
ANOVAsum
ANOVAplot <- ggplot(data=ANOVAdata,aes(x=TGroup,y=resVar,fill=TGroup)) +
  geom_boxplot() +
  xlab("elevational group") +
  ylab("leaf density")
print(ANOVAplot)
```
After running it several times, I observe that the results are not that different with each run.  

##### Question 6: Adjust the means & determine the effect size (aka how small the difference can be b/n the groups to still get a p-value less than 0.05    
```{r}
nGroup <- 3 
nName <- c("low","mid","high")
nSize <- c(45,45,45) 
nMean <- c(140,150,160)# I adjusted the difference b/n the means to be much smaller
nSD <- c(10,10,10) 
ID <- 1:(sum(nSize)) 
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]))
TGroup <- factor(x=rep(nName,nSize), levels=nName, ordered= is.ordered(nName)) 
ANOVAdata <- data.frame(ID,TGroup,resVar) 
ANOVAmodel <- aov(resVar~TGroup,data=ANOVAdata) 
boop <- summary(ANOVAmodel) 
aggregate(resVar~TGroup,data=ANOVAdata,FUN=mean)
unlist(boop)
ANOVAsum <- list(Fval=unlist(boop)[7],probF=unlist(boop)[9])
ANOVAsum
ANOVAplot <- ggplot(data=ANOVAdata,aes(x=TGroup,y=resVar,fill=TGroup)) +
  geom_boxplot() +
  xlab("elevational group") +
  ylab("leaf density")
print(ANOVAplot)
# that one was still very significantly different, so I'll try again

nGroup <- 3 
nName <- c("low","mid","high")
nSize <- c(45,45,45) 
nMean <- c(145,150,155)# going even smaller this time
nSD <- c(10,10,10) 
ID <- 1:(sum(nSize)) 
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]))
TGroup <- factor(x=rep(nName,nSize), levels=nName, ordered= is.ordered(nName)) 
ANOVAdata <- data.frame(ID,TGroup,resVar) 
ANOVAmodel <- aov(resVar~TGroup,data=ANOVAdata) 
boop <- summary(ANOVAmodel) 
aggregate(resVar~TGroup,data=ANOVAdata,FUN=mean)
unlist(boop)
ANOVAsum <- list(Fval=unlist(boop)[7],probF=unlist(boop)[9])
ANOVAsum
ANOVAplot <- ggplot(data=ANOVAdata,aes(x=TGroup,y=resVar,fill=TGroup)) +
  geom_boxplot() +
  xlab("elevational group") +
  ylab("leaf density")
print(ANOVAplot)
#still significant

nGroup <- 3 
nName <- c("low","mid","high")
nSize <- c(45,45,45) 
nMean <- c(149,150,151)# smaller!
nSD <- c(10,10,10) 
ID <- 1:(sum(nSize)) 
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]))
TGroup <- factor(x=rep(nName,nSize), levels=nName, ordered= is.ordered(nName)) 
ANOVAdata <- data.frame(ID,TGroup,resVar) 
ANOVAmodel <- aov(resVar~TGroup,data=ANOVAdata) 
boop <- summary(ANOVAmodel) 
aggregate(resVar~TGroup,data=ANOVAdata,FUN=mean)
unlist(boop)
ANOVAsum <- list(Fval=unlist(boop)[7],probF=unlist(boop)[9])
ANOVAsum
ANOVAplot <- ggplot(data=ANOVAdata,aes(x=TGroup,y=resVar,fill=TGroup)) +
  geom_boxplot() +
  xlab("elevational group") +
  ylab("leaf density")
print(ANOVAplot)
# this one wasn't significantly different
# so having a difference of 5 g/mm^2 b/n groups was significant, but 1 g/mm^2 is not
# let's try 3 now

nGroup <- 3 
nName <- c("low","mid","high")
nSize <- c(45,45,45) 
nMean <- c(147,150,153)# smaller!
nSD <- c(10,10,10) 
ID <- 1:(sum(nSize)) 
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]))
TGroup <- factor(x=rep(nName,nSize), levels=nName, ordered= is.ordered(nName)) 
ANOVAdata <- data.frame(ID,TGroup,resVar) 
ANOVAmodel <- aov(resVar~TGroup,data=ANOVAdata) 
boop <- summary(ANOVAmodel) 
aggregate(resVar~TGroup,data=ANOVAdata,FUN=mean)
unlist(boop)
ANOVAsum <- list(Fval=unlist(boop)[7],probF=unlist(boop)[9])
ANOVAsum
ANOVAplot <- ggplot(data=ANOVAdata,aes(x=TGroup,y=resVar,fill=TGroup)) +
  geom_boxplot() +
  xlab("elevational group") +
  ylab("leaf density")
print(ANOVAplot)
#this one was not significant! p = 0.054
```
The effect size is approximately 4 g/mm^2. When there is a difference of 4 g/mm^2 b/n the means, the results are usually statistically significant (p<0.05), but when the difference is 3, they're usually not significant.  


##### Question 7: For the effect sizes I hypothesized (50 g/mm^2 b/n groups), what is the minimum sample size needed to detect a statistically significant effect?  
```{r}
nGroup <- 3 
nName <- c("low","mid","high")
nSize <- c(2,2,2) 
nMean <- c(100,150,200)
nSD <- c(10,10,10) 
ID <- 1:(sum(nSize)) 
resVar <- c(rnorm(n=nSize[1],mean=nMean[1],sd=nSD[1]),
            rnorm(n=nSize[2],mean=nMean[2],sd=nSD[2]),
            rnorm(n=nSize[3],mean=nMean[3],sd=nSD[3]))
TGroup <- factor(x=rep(nName,nSize), levels=nName, ordered= is.ordered(nName)) 
ANOVAdata <- data.frame(ID,TGroup,resVar) 
ANOVAmodel <- aov(resVar~TGroup,data=ANOVAdata) 
boop <- summary(ANOVAmodel) 
aggregate(resVar~TGroup,data=ANOVAdata,FUN=mean)
unlist(boop)
ANOVAsum <- list(Fval=unlist(boop)[7],probF=unlist(boop)[9])
ANOVAsum
ANOVAplot <- ggplot(data=ANOVAdata,aes(x=TGroup,y=resVar,fill=TGroup)) +
  geom_boxplot() +
  xlab("elevational group") +
  ylab("leaf density")
print(ANOVAplot)
```
It seems that given the standard deviation, with any reasonable sample size (greater than 1), my expected means lead to significant results.  
Perhaps my hypothesized effect size is too large?




